---
title: "Week02PC_Gapminder"
author: "Andrew Croft"
date: "07/03/2021"
output:
  pdf_document:
    df_print: paged
---

```{r setup, include=FALSE}
#clear workspace
a<-1
rm(list=ls())
# if you need current working directory, use getwd()
# set working directory, use fwd slash not back.
setwd("C:/Users/cacro/Documents/R/Data_Analysis_Actuaries/Week02PC_Gapminder")
#
knitr::opts_chunk$set(echo = TRUE, fig.pos='h')
library(tidyverse)
library(gapminder)
data(gapminder)
```

## Introduction

In this document, we will be exploring the Gapminder dataset.  And I hate git push.

## Observations of Gapminder

### Countries with life expectancy less than 30 at any time in dataset

```{r, echo=FALSE}
lowlc<- gapminder %>% filter(lifeExp<30)
knitr::kable(lowlc,format="markdown")
```

### Afghanistan's Life Expectancy over time
**Rose strongly from 1952 but plateaued in 1980s with 30 years of war between the Russians, Taliban, Americans plus allies including Australia. Amazingly heading up again at 2007.**

```{r, echo=FALSE}
data01 <- gapminder %>% filter(country=="Afghanistan") %>% select(year,lifeExp)
plot(data01$year, data01$lifeExp,type="b")
```

### Rwanda's Life Expectancy over time
**Calamity in 1992 - war as I recall.**

```{r Rwanda Life Expectancy over Time, echo=FALSE}
data02 <- gapminder %>% filter(country=="Rwanda") %>% select(year,lifeExp)
plot(data02$year, data02$lifeExp,type="b")
```


### Countries with life expectancy greater than 81

```{r, echo=FALSE}
highlc<- gapminder %>% filter(lifeExp>81) %>% arrange(desc(lifeExp))
knitr::kable(highlc,format="markdown")
```

### Japan's Life Expectancy over time
**Positive trend post-war, but also reflects their post war economic 'miracle'.**

```{r Japan Life Expectancy over Time, echo=FALSE}
data01 <- gapminder %>% filter(country=="Japan") %>% select(year,lifeExp)
plot(data01$year, data01$lifeExp,type="b")
```

### Hong Kong, China's Life Expectancy over time

```{r Hong Kong Life Expectancy over Time, echo=FALSE}
data02 <- gapminder %>% filter(country=="Hong Kong, China") %>% select(year,lifeExp)
plot(data02$year, data02$lifeExp,type="b")
```

### Iceland's Life Expectancy over time

```{r Iceland Life Expectancy over Time, echo=FALSE}
data02 <- gapminder %>% filter(country=="Iceland") %>% select(year,lifeExp)
plot(data02$year, data02$lifeExp,type="b")
```

### Switzerland's Life Expectancy over time

```{r Switzerland Life Expectancy over Time, echo=FALSE}
data02 <- gapminder %>% filter(country=="Switzerland") %>% select(year,lifeExp)
plot(data02$year, data02$lifeExp,type="b")
```

### Australia's Life Expectancy over time
**Consistent with other developed countries, a positive trend since 1952.**

```{r Australia Life Expectancy over Time, echo=FALSE}
data02 <- gapminder %>% filter(country=="Australia") %>% select(year,lifeExp)
plot(data02$year, data02$lifeExp,type="b")
```


### Life expectancy by continent in 2007

*How good is it to be in Oceania?  Hmmm....*

```{r, echo=FALSE}
data03 <-gapminder %>% filter(year==2007) 
boxplot(lifeExp~continent,data=data03)
```
```{r, echo=FALSE}
# Script run completed.
```

